---
title: "ARMA estimation"
author: "Giovani Valdrighi, Vitória Guardieiro"
date: "05/11/2020"
output: beamer_presentation
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tseries)
library(fpp2)
library(EnvStats)
library(zoo)
library(car)
```

# Modelling process

- Plot the data and look for patterns.
- If necessary, remove seasonality.
- If necessary, use BoxCox transformation to stabilize the variance.
- If necessary, difference the data until is stationary. Use of Dickey Fuller test.
- Plot the ACF and PACF to identify the model order, $p$ and $q$ for $ARMA(p, q)$.
- Compare identified models, chose the one that minimize the AIC
- Analysis of the residuals of the model, with ACF and histogram.
- If the residuals look like white noise, make forecasts.

# WMurders

```{r}
data <- wmurders
par(mfrow = c(1, 1))
plot(data, 
     main = "Monthly female murder rate \n(per 100,000 standard population)", 
     ylab = "Rate")
```

--- 

## Variance stabilization and stationarity

- There is no need to stabilize data, we can look at the tendency with Dick-Fulley.
```{r}
adf.test(data)
```
- Data is not stationary, we are going to test with one and two differences.

--- 

```{r}
adf.test(diff(data))
```

```{r}
adf.test(diff(data, differences = 2))
```
- Diff 1 P-value: 0.02726

- Diff 2 P-value: less than 0.01 <- choosen. 

---

```{r}
plot(diff(data, differences = 2), 
     ylab = 'Value', 
     main = "Monthly female murder rate (diff 2) \n(per 100,000 standard population)", )
```

---

## ACF and PACF

```{r}
par(mfrow = c(2, 1))
acf(diff(data, differences = 2),
    main = "ACF Monthly female murder rate (diff 2)")
pacf(diff(data, differences = 2),
     main = "PACF   Monthly female murder rate (diff 2)")
```

- Looks like AR(1) because the first spike of the PACF. The models that will be tested are AR(1), MA(1), ARMA(1, 1).

---

## AR(1)

```{r}
model.AR1 <- Arima(data, order = c(1, 2,0 ), include.mean = FALSE)
summary(model.AR1)
```

---

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.AR1, h = 3))
lines(model.AR1$fitted, col = 'red')
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```


---

### Residuals analysis

```{r}
checkresiduals(model.AR1, test = FALSE)
```

---

```{r}
qqPlot(model.AR1$residuals, main = "AR(1) model residuals")
```

---

```{r}
jarque.bera.test(model.AR1$residuals)
```

---

## MA(1)

```{r}
model.MA1 <- Arima(data, order = c(0, 2,1 ), include.mean = FALSE)
summary(model.MA1)
```

---

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.MA1, h = 3))
lines(model.MA1$fitted, col = 'red')
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```

---

### Residuals analysis

```{r}
checkresiduals(model.MA1, test = FALSE)
```

---

```{r}
qqPlot(model.MA1$residuals, main = "MA(1) model residuals")
```

---

```{r}
jarque.bera.test(model.MA1$residuals)
```

---

## ARMA(1, 1)


```{r}
model.ARMA11 <- Arima(data, order = c(1, 2, 1 ), include.mean = FALSE)
summary(model.ARMA11)
```

---

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.ARMA11, h = 3))
lines(model.ARMA11$fitted, col = 'red')
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```


---

### Residuals analysis

```{r}
checkresiduals(model.ARMA11, test = FALSE)
```

---

```{r}
qqPlot(model.ARMA11$residuals, main = "AR(1) model residuals")
```

---

```{r}
jarque.bera.test(model.ARMA11$residuals)
```

---

## Model selection

- The model with lowest AIC is the $ARMA(1, 1)$. The final forecast is:

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.ARMA11, h = 3))
lines(model.ARMA11$fitted, col = 'red')
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```

# Debit Cards

```{r}
data <- debitcards
plot(data, main = 'Retail Debit Card Usage In Iceland')
```

---

## Stabilize variance

```{r}
lambda <- BoxCox.lambda(data)
data.stabilized <- BoxCox(data, lambda)
plot(data.stabilized, 
     main = 'Retail Debit Card Usage In Iceland after BoxCox', 
     ylab = 'Value')
```

---

## Remove seasonality

- Removing with STL (Lowess fit).

```{r}
stl.fit <- stl(data.stabilized, s.window = 12)
plot(stl.fit)
```

---

```{r}
data.trend <- data.stabilized - stl.fit$time.series[, 1]
plot(data.trend, main = "Debit Card Usage - \n Stabilized and without seasonality", 
     ylab = "Value")
```

---

## Stationarity

- Dickey Fuller test with series and differenced series.

```{r}
adf.test(data.trend)
#P-value is not lower than 0.05

#Testing diff 1
adf.test(diff(data.trend))
```

---

```{r}
plot(diff(data.trend), 
     main = 'DebitCard Series (diff 1)', 
     ylab = 'Value')
```

---

## ACF and PACF

```{r}
par(mfrow = c(2, 1))
acf(diff(data.trend), main = "ACF DebitCard Series (diff 1)")
pacf(diff(data.trend), main = "PACF DebitCard Series (diff 1)")
```

- Looks like AR(2) because of the first two lags. Models being tested are AR(1), AR(2) and ARMA(2, 1). The data is periodic with one year, because of that we are going to use a window of 2 years to predict 3 months.

---

## AR(1)

```{r}
eval_AR1 <- function(data){
  m <- Arima(data, order = c(1, 1, 0))
  return(AIC(m))
}
AIC_AR1 <- rollapply(data.trend, width = 24, FUN = eval_AR1)
```

- AR(1) forecast in the last window.

```{r}
model.AR1 <- Arima(data.trend[141:164], order = c(1, 1, 0))
summary(model.AR1)
```

--- 

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.AR1, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(20, 25))
lines(InvBoxCox(model.AR1$fitted, lambda = lambda), col = 'red')
lines(InvBoxCox(data.trend[141:164], lambda = lambda))
```

---

## AR(2)

- AR(2) forecast in the last window.

```{r}
eval_AR2 <- function(data){
  m <- Arima(data, order = c(2, 1, 0), lambda = lambda)
  return(AIC(m))
}
AIC_AR2 <- rollapply(data.trend, width = 24, FUN = eval_AR2)

model.AR2 <- Arima(data.trend[141:164], order = c(2, 1, 0))
summary(model.AR2)
```

---


```{r}
plot(forecast(model.AR2, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(20, 25))
lines(InvBoxCox(model.AR2$fitted, lambda = lambda), col = 'red')
lines(InvBoxCox(data.trend[141:164], lambda = lambda))
```

---

## ARMA(2, 1)

- Forecast in the last window.

```{r}
eval_ARMA21 <- function(data){
  m <- Arima(data, order = c(2, 1, 1), lambda = lambda)
  return(AIC(m))
}
AIC_ARMA21 <- rollapply(data.trend, width = 24, FUN = eval_ARMA21)

model.ARMA21 <- Arima(data.trend[141:164], order = c(2, 1, 1))
summary(model.ARMA21)
```

---

```{r}
plot(forecast(model.ARMA21, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(20, 25))
lines(InvBoxCox(model.ARMA21$fitted, lambda = lambda), col = 'red')
lines(InvBoxCox(data.trend[141:164], lambda = lambda))

```

--- 

- Comparison between AIC of the three models in each of the windows.

```{r}
plot(AIC_AR1, ylim = c(-150, -50), ylab = "AIC", main = "Comparison between AICs of models")
lines(AIC_AR2, col = "blue")
lines(AIC_ARMA21, col = "red")
legend("bottomleft", legend=c("AR(1)", "AR(2)", "ARMA(2, 1)"), col=c('black','blue', 'red'), lty = 1:1, cex=0.8)

```

- The best model is AR(2).

---

## Residual analysis

- Analysis in the last window.

```{r}
checkresiduals(model.AR2, lag = 24)
```

---

```{r}
jarque.bera.test(model.AR2$residuals)
```

---

## Model selection

- Final model is AR(2). We are going to make predictions on the original scale, before BoxCox and STL.

```{r}
plot(forecast(model.AR2, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(10, 30))
lines(InvBoxCox(model.AR2$fitted, lambda = lambda), col = 'red')
lines(data[141:164])
lines(InvBoxCox(data.trend[141:164], lambda = lambda), col = "blue")
legend("bottomleft", legend=c("Real", "Real without seasonality", "Prediction without seasonality"), col=c('black','blue', 'red'), lty = 1:1, cex=0.8)

```


---

```{r}
season.forecast <- stl.fit$time.series[(165-12):(167-12), 1]
plot(forecast(model.AR2, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(10, 30))
lines(InvBoxCox(model.AR2$fitted + stl.fit$time.series[141:164, 1], lambda = lambda), col = 'red')
lines(data[141:164])
legend("bottomleft", legend=c("Real", "Real without seasonality", "Prediction without seasonality"), col=c('black','blue', 'red'), lty = 1:1, cex=0.8)

```

# Eating Out

```{r}
data <- auscafe
plot(data, 
     main = "Monthly expenditure on eating out in Australia", 
     ylab = "Expenditure")
```

---

## Stabilize variance

```{r}
lambda <- BoxCox.lambda(data)
data.stabilized <- BoxCox(data, lambda)
plot(data.stabilized, 
     main = 'Monthly expenditure on eating out in Australia after BoxCox', 
     ylab = 'Value')
```

---

## Remove seasonality

- Removing with STL (Lowess fit).

```{r}
stl.fit <- stl(data.stabilized, s.window = 12)
plot(stl.fit)
```

---

```{r}
data.trend <- data.stabilized - stl.fit$time.series[, 1]
plot(data.trend, main = "Monthly expenditure on eating out - \n Stabilized and without seasonality", 
     ylab = "Value")
```

---

## Stationarity

- Dickey Fuller test with series and differenced series.

```{r}
adf.test(data.trend)
#P-value is not lower than 0.05

#Testing diff 1
adf.test(diff(data.trend))
```

---

```{r}
plot(diff(data.trend), 
     main = 'auscafe series (diff 1)', 
     ylab = 'Value')
```

---

## ACF and PACF

```{r}
par(mfrow = c(2, 1))
acf(diff(data.trend), main = "ACF auscafe series (diff 1)")
pacf(diff(data.trend), main = "PACF auscafe series (diff 1)")
```

- Looks like AR(1) because of the first two lags. Models being tested are AR(1), AR(2), ARMA(1,1) and ARMA(2, 1). The data is periodic with one year, because of that we are going to use a window of 2 years to predict 3 months.

---

## AR(1)

```{r}
eval_AR1 <- function(data){
  m <- Arima(data, order = c(1, 1, 0))
  return(AIC(m))
}
AIC_AR1 <- rollapply(data.trend, width = 24, FUN = eval_AR1)
```

- AR(1) forecast in the last window.

```{r}
model.AR1 <- Arima(data.trend[403:426], order = c(1, 1, 0))
summary(model.AR1)
```

--- 

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.AR1, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(3.4, 3.8))
lines(InvBoxCox(model.AR1$fitted, lambda = lambda), col = 'red')
lines(InvBoxCox(data.trend[403:426], lambda = lambda))
```

---

## AR(2)

- AR(2) forecast in the last window.

```{r}
eval_AR2 <- function(data){
  m <- Arima(data, order = c(2, 1, 0), lambda = lambda, method="ML")
  return(AIC(m))
}
AIC_AR2 <- rollapply(data.trend, width = 24, FUN = eval_AR2)

model.AR2 <- Arima(data.trend[403:426], order = c(2, 1, 0))
summary(model.AR2)
```

---


```{r}
plot(forecast(model.AR2, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(3.4, 3.8))
lines(InvBoxCox(model.AR2$fitted, lambda = lambda), col = 'red')
lines(InvBoxCox(data.trend[403:426], lambda = lambda))
```

---

## ARMA(1, 1)

- Forecast in the last window.

```{r}
eval_ARMA11 <- function(data){
  m <- Arima(data, order = c(1, 1, 1), lambda = lambda, method="ML")
  return(AIC(m))
}
AIC_ARMA11 <- rollapply(data.trend, width = 24, FUN = eval_ARMA21)

model.ARMA11 <- Arima(data.trend[403:426], order = c(1, 1, 1))
summary(model.ARMA11)
```

---

```{r}
plot(forecast(model.ARMA11, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(3.4, 3.8))
lines(InvBoxCox(model.ARMA11$fitted, lambda = lambda), col = 'red')
lines(InvBoxCox(data.trend[403:426], lambda = lambda))

```

---

## ARMA(2, 1)

- Forecast in the last window.

```{r}
eval_ARMA21 <- function(data){
  m <- Arima(data, order = c(2, 1, 1), lambda = lambda, method="ML")
  return(AIC(m))
}
AIC_ARMA21 <- rollapply(data.trend, width = 24, FUN = eval_ARMA21)

model.ARMA21 <- Arima(data.trend[403:426], order = c(2, 1, 1))
summary(model.ARMA21)
```

---

```{r}
plot(forecast(model.ARMA21, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(3.4, 3.8))
lines(InvBoxCox(model.ARMA21$fitted, lambda = lambda), col = 'red')
lines(InvBoxCox(data.trend[403:426], lambda = lambda))

```

--- 

- Comparison between AIC of the three models in each of the windows.

```{r}
plot(AIC_AR1, ylim = c(-150, 150), ylab = "AIC", main = "Comparison between AICs of models")
lines(AIC_AR2, col = "blue")
lines(AIC_ARMA11, col = "green")
lines(AIC_ARMA21, col = "red")
legend("topleft", legend=c("AR(1)", "AR(2)", "ARMA(1, 1)", "ARMA(2, 1)"), col=c('black','blue', 'green','red'), lty = 1:1, cex=0.8)

```

- The best model is AR(1).

---

## Residual analysis

- Analysis in the last window.

```{r}
checkresiduals(model.AR1, lag = 24)
```

---

```{r}
jarque.bera.test(model.AR1$residuals)
```

---

## Model selection

- Final model is AR(1). We are going to make predictions on the original scale, before BoxCox and STL.

```{r}
plot(forecast(model.AR1, h = 3, lambda = lambda, biasadj = TRUE), include = 0, 
     xlim = c(0, 28), ylim = c(3.2, 4.1))
lines(InvBoxCox(model.AR1$fitted, lambda = lambda), col = 'red')
lines(data[403:426])
lines(InvBoxCox(data.trend[403:426], lambda = lambda), col = "blue")
legend("topleft", legend=c("Real", "Real without seasonality", "Prediction without seasonality"), col=c('black','blue', 'red'), lty = 1:1, cex=0.8)

```


