---
title: "ARMA estimation"
author: "Giovani Valdrighi, Vitória Guardieiro"
date: "05/11/2020"
output: beamer_presentation
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tseries)
library(fpp2)
library(EnvStats)
library(zoo)
library(car)
```

# Modelling process

- Plot the data and look for patters.
- If necessary, use BoxCox transformation to stabilize the variance.
- If necessary, remove seasonality.
- If necessary, difference the data until is stationary. Use of Dickey Fuller test.
- Plot the ACF and PACF to identify the model order, $p$ and $q$ for $ARMA(p, q)$.
- Compare identified models, chose the one that minimize the AIC
- Analysis of the residuals of the model, with ACF and histogram.
- If the residuals look like white noise, make forecasts.

# WMurders

```{r}
data <- wmurders
par(mfrow = c(1, 1))
plot(data, 
     main = "Monthly female murder rate \n(per 100,000 standard population)", 
     ylab = "Rate")
```

--- 

## Variance stabilization and stationarity

- There is no need to stabilize data, we can look at the tendency with Dick-Fulley.
```{r}
adf.test(data)
```
- Data is not stationary, we are going to test with one and two differences.

--- 

```{r}
adf.test(diff(data))
```

```{r}
adf.test(diff(data, differences = 2))
```
- Diff 1 P-value: 0.02726

- Diff 2 P-value: less than 0.01 <- choosen. 

---

```{r}
plot(diff(data, differences = 2), 
     ylab = 'Value', 
     main = "Monthly female murder rate (diff 2) \n(per 100,000 standard population)", )
```

---

## ACF and PACF

```{r}
par(mfrow = c(2, 1))
acf(diff(data, differences = 2),
    main = "ACF Monthly female murder rate (diff 2)")
pacf(diff(data, differences = 2),
     main = "PACF   Monthly female murder rate (diff 2)")
```

- Looks like AR(1) because the first spike of the PACF. The models that will be tested are AR(1), MA(1), ARMA(1, 1).

---

## AR(1)

```{r}
model.AR1 <- Arima(data, order = c(1, 2,0 ), include.mean = FALSE)
summary(model.AR1)
```

---

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.AR1, h = 3))
lines(model.AR1$fitted, col = 'red')
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```


---

### Residuals analysis

```{r}
checkresiduals(model.AR1, test = FALSE)
```

---

```{r}
qqPlot(model.AR1$residuals, main = "AR(1) model residuals")
```

---

```{r}
jarque.bera.test(model.AR1$residuals)
```

---

## MA(1)

```{r}
model.MA1 <- Arima(data, order = c(0, 2,1 ), include.mean = FALSE)
summary(model.MA1)
```

---

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.MA1, h = 3))
lines(model.MA1$fitted, col = 'red')
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```

---

### Residuals analysis

```{r}
checkresiduals(model.MA1, test = FALSE)
```

---

```{r}
qqPlot(model.MA1$residuals, main = "MA(1) model residuals")
```

---

```{r}
jarque.bera.test(model.MA1$residuals)
```

---

## ARMA(1, 1)


```{r}
model.ARMA11 <- Arima(data, order = c(1, 2, 1 ), include.mean = FALSE)
summary(model.ARMA11)
```

---

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.ARMA11, h = 3))
lines(model.ARMA11$fitted, col = 'red')
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```


---

### Residuals analysis

```{r}
checkresiduals(model.ARMA11, test = FALSE)
```

---

```{r}
qqPlot(model.ARMA11$residuals, main = "AR(1) model residuals")
```

---

```{r}
jarque.bera.test(model.ARMA11$residuals)
```

---

## Model selection

- The model with lowest AIC is the $ARMA(1, 1)$. The final forecast is:

```{r}
par(mfrow = c(1, 1))
plot(forecast(model.ARMA11, h = 3))
lines(model.ARMA11$fitted, col = 'red')
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```
