---
title: "ARIMA Identification"
author: "Giovani Valdrighi, Vitória Guardieiro"
date: "14/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tseries)
library(forecast)
library(car)
```

# Simulation of the distribution of the Dickey Fuller test statistic 

```{r}
TT <-  100
n <- 50
S <- 10000
phi <- 0.8
```

```{r}
alpha_f <- function(y){
  ax1 <- 0
  ax2 <- 0 
  for(i in 1:(length(y)-1)){
    ax1 <- ax1 + y[i]*y[i+1]
    ax2 <- ax2 + y[i]*y[i]
  }
  alpha <- ax1/ax2 - 0.2
  return(alpha)
}

s_f <- function(y, alpha){
  axs <- 0
  ax <- 0
  for(i in  1:(length(y)-1)){
    axs <- axs + (y[i+1]-y[i]-alpha*y[i])^2
    ax <- ax + y[i]^2
  }
  axs <- axs/(length(y)-1)
  ax <- ax^0.5
  s <- axs/ax
  return(s)
}

t_f <- function(y){
  alp <- alpha_f(y)
  s <- s_f(y, alp)
  t <- alp/s
  return(t)
}
```

```{r}
axi <- 1
all_t <- list()

for(j in 1:S){
  wn <- rnorm(TT+n)
  y <- rep(0, TT+n)
  for(i in 1:(TT+n-1)){
    y[i+1] <- phi*y[i]+wn[i]
  }

  for(i in 1:n){
    all_t[[axi]] <- t_f(y[i:TT+i])
    axi <- axi+1
  }
}
```

```{r}
t_num <- unlist(all_t)
hist(t_num)
```

### 2.3 Intercept

```{r}
interc <- 10

axi <- 1
all_t <- list()

for(j in 1:S){
  wn <- rnorm(TT+n)
  y <- rep(0, TT+n)
  y[1] <- interc
  for(i in 1:(TT+n-1)){
    y[i+1] <- interc+phi*y[i]+wn[i]
  }

  for(i in 1:n){
    all_t[[axi]] <- t_f(y[i:TT+i])
    axi <- axi+1
  }
}
```

```{r}
t_num <- unlist(all_t)
hist(t_num)
```



# ARIMA Identification with IPCA data

```{r}
ipca <- read.csv("ipca.csv")
ipca.ts <- ts(ipca$ipca.acum, start = c(1994, 8), frequency = 12)
plot(ipca.ts, type = "l", main = "IPCA by month", ylab = "IPCA", xlab = "Date")
```

Steps for the ARIMA model identification:

  - Verify if the series need a transformation to estabilize the variance.
  - Differentiate the series how many times it's needed to the series ACF decay fast to 0. In this step it can also be used the Dickey Fuller test.
  - Indentify the ARMA(p, q) process by the ACF and PACF plots.

We see that our data show a small variance, and it's the same along the observed period. Now we differentiate the serie.

```{r}
diff_series <- function(data, d){
  if(d > 0){
    for(i in 1:d){
      data = diff(data)
    }
  }
  par(mfrow = c(1, 2))
  acf(data, main = "IPCA ACF")
  pacf(data, main = "IPCA PACF")
}
```

### Diff 0 times

```{r}
diff_series(ipca.ts, 0)
```

### Diff 1 time

```{r}
diff_series(ipca.ts, 1)
```
```{r}
plot((diff(ipca.ts)), main = "IPCA by month (diff 1)", xlab = "Date", ylab = "IPCA (diff 1)")
```



### Diff 2 times

```{r}
diff_series(ipca.ts, 2)
```

```{r}
plot(diff(diff(ipca.ts)), main = "IPCA by month (diff 2)", xlab = "Date", ylab = "IPCA (diff 2)")
```

With 2 differentiation the ACF decay fast, but the first lag is negative on the ACF, what is a indicator of over-differenced. So let's apply the Dickey Fuller test on both models.

```{r}
diff1_ipca <- diff(ipca.ts, differences = 1)
diff2_ipca <- diff(ipca.ts, differences = 2)
adf.test(diff1_ipca)
```

```{r}
adf.test(diff2_ipca)
```

Let's keep only one differentiation.

### Identify ARMA(p, q)

Now let's look at the IPCA to identify the ARMA(p, q) order.

```{r}
diff_series(diff1_ipca, d = 0)
```

As the PACF decays fast, it appears to be a MA model, and the signicant values for the ACF are before the lag 4, so we start with testing the model MA(4).

```{r}
model.ma <- arma(diff1_ipca, order = c(0, 4)) 
summary(model.ma)
plot(diff1_ipca, ylab = "IPCA (diff 1)", xlab = "Date", main = "IPCA (diff 1) MA(4)")
lines(model.ma$fitted.values, col = "red")
```

We can use the function _auto.arima_ that try different models and choose the with the best AIC.

```{r}
model.auto <- auto.arima(diff1_ipca, ic = "aic")
summary(model.auto)
plot(diff1_ipca, ylab = "IPCA (diff 1)", xlab = "Date", main = "IPCA (diff 1) ARIMA(0, 1, 3)")
lines(model.auto$fitted, col = "red")
```

We see that the auto ARIMA function fits an ARIMA(0, 1, 3), but the AIC is bigger than our MA(4).

### MA(4)

Let's look at the residuals from our MA(4) model.

```{r}
qqPlot(model.ma$residuals, main = "QQPlot of residuals")
```

```{r}
par(mfrow = c(1, 2))
acf(model.ma$residuals, na.action = na.pass, main = "ACF of residuals")
pacf(model.ma$residuals, na.action = na.pass, main = "PACF of residuals")
```

