---
title: "Ozonio model"
author: "Giovani Valdrighi, Vitória Guardieiro"
date: "24/09/2020"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

## Data

- New York data from 15/07/2001 to 30/04/2016.

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(VIM)
library(zoo)
data <- read.csv("NY.csv")
data <- data %>% select(Date.Local, O3.Mean) %>% mutate(Date.Local = as.Date(Date.Local))
data <- data[-1, ]
startDate <- data$Date.Local[1] #First observed sunday
finalDate <- data$Date.Local[length(data$Date.Local)] #Last observed day
days <- seq(from = startDate, to = finalDate, by = 1)
dayWeek <- seq(from = startDate, to =  finalDate, by = 7)
#fulldata have NA rows in days that there is no observation
fulldata <- data.frame(Date.Local = days)
fulldata <- as.data.frame(merge(data, fulldata, all.y = TRUE)) %>% mutate(Date.Local = as.Date(Date.Local))
ggplot(data = fulldata) +
  geom_point(aes(x = Date.Local, y = O3.Mean)) +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y", name = "Date") +
  scale_y_continuous(name = "O3") +
  ggtitle("Daily O3 mean") + 
  theme_bw()
```

## Missing data

- There are 52 time skips in the data, in a total of 473 days.
- The biggest skips is 108 days in 2011.
- The majority of skips are of 1 or 2 days.
- Around 9.5% missing data.
- The missing observations are distributed along the time without a clear pattern.


---

```{r}
dateJumps <- data$Date.Local - lag(data$Date.Local, n = 1)
dateJumps[is.na(dateJumps)] <- FALSE
ggplot(data = data[dateJumps > 1, ]) + 
  geom_point(aes(x = Date.Local, y = O3.Mean), colour = 'red') +
  scale_x_date(name = "Date") + 
  scale_y_continuous(name = "O3") +
  ggtitle("Observations after data skips") +
  theme_bw()
```

----

```{r}
skips <- dateJumps[dateJumps > 1]
ggplot(data = as.data.frame(skips)) +
  geom_bar(aes(skips), fill = "blue") +
  scale_y_continuous(name = "Count") +
  scale_x_continuous(name = "Quantity of days", breaks = seq(0, 150, 5)) +
  ggtitle("Days missing in sequence") +
  theme_bw()
```

## Imputation method

- It was used the kNN method to imputate values on missing observations.
- The kNN method needs the parameter k, the number of closest points considered.
- Starting with k = 7.

---

```{r}
imputedData <- data.frame("obs" = index(fulldata), "O3.Mean" = fulldata$O3.Mean)
imputedData <- kNN(imputedData, variable = "O3.Mean", k = 7) #K = 7 because we'll consider the week of the missing data
imputedData$Date.Local <- days

ggplot() +
  geom_point(data = imputedData[is.na(fulldata$O3.Mean), ], aes(x = Date.Local, y = O3.Mean), colour = 'red') +
  geom_point(data = fulldata, aes(x = Date.Local, y = O3.Mean), alpha = 0.2, fill = "black") +
  scale_y_continuous(name = "O3") +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y", name = "Date") +
  ggtitle("Real vs Imputed data") +
  theme_bw()
```

---

- Method create a bad behavior where the size of the skips is bigger than 7 days.

```{r}
ggplot() +
  geom_point(data = imputedData[is.na(fulldata$O3.Mean), ], aes(x = Date.Local, y = O3.Mean), colour = 'red') +
  geom_point(data = fulldata, aes(x = Date.Local, y = O3.Mean), alpha = 0.2, fill = "black") +
  scale_y_continuous(name = "O3") +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y", name = "Date", limits = c(as.Date("2011-01-01"), as.Date("2012-01-01"))) +
  ggtitle("Real vs Imputed data") +
  theme_bw()
```


---

- To deal with this, the parameter k used for imputation will be different if the size of the skip is minor tem 30 days, between 30 days and 100 days, or bigger than 100 days.
- k = 7, k = 45, k = 120, respectively.
- We will aggregate closest points by weighted by distance mean.

---

```{r}
imputedData[((is.na(fulldata$O3.Mean)) 
             & (fulldata$Date.Local > "2003-08-01") 
             & (fulldata$Date.Local < "2003-09-15")),"O3.Mean"] <- NA

imputedData[((is.na(fulldata$O3.Mean)) 
             & (fulldata$Date.Local > "2009-07-01") 
             & (fulldata$Date.Local < "2009-08-20")),"O3.Mean"] <- NA

imputedData <- kNN(imputedData, variable = c("O3.Mean"), numFun = weighted.mean, k = 45, weightDist = TRUE)


imputedData[((is.na(fulldata$O3.Mean)) 
              & (fulldata$Date.Local > "2011-01-01") 
              & (fulldata$Date.Local < "2012-01-01")),"O3.Mean"] <- NA

imputedData <- kNN(imputedData, variable = c("O3.Mean"), numFun = weighted.mean, k = 120, weightDist = TRUE)

ggplot() +
  geom_point(data = imputedData[is.na(fulldata$O3.Mean), ], aes(x = Date.Local, y = O3.Mean), colour = 'red') +
  geom_point(data = fulldata, aes(x = Date.Local, y = O3.Mean), alpha = 0.2, fill = "black") +
  scale_y_continuous(name = "O3") +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y", name = "Date") +
  ggtitle("Real vs Imputed data (by separeted imput.)") +
  theme_bw()

```

---

```{r}
ggplot() +
  geom_point(data = imputedData[is.na(fulldata$O3.Mean), ], aes(x = Date.Local, y = O3.Mean), colour = 'red') +
  geom_point(data = fulldata, aes(x = Date.Local, y = O3.Mean), alpha = 0.2, fill = "black") +
  scale_y_continuous(name = "O3") +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y", name = "Date", limits = c(as.Date("2011-01-01"), as.Date("2012-01-01"))) +
  ggtitle("Real vs Imputed data (by separated imput.)") +
  theme_bw()
```

## Weekly data
```{r}
k = 0
val = 1
l = rep(NA, length(fulldata$Date.Local) - 1)
while(k < length(fulldata$Date.Local) - 1){
  for(i in 1:7){
    l[k+i] = val
  }
  val = val + 1
  k = k +7
}

weekData <- cbind(fulldata)
weekData$weekInd <- l
weekData <- weekData %>% group_by(weekInd) %>% summarise(O3.Mean = mean(O3.Mean, na.rm = TRUE))
weekData$Date.Local <- dayWeek
ggplot(weekData) +
  geom_point(aes(x = Date.Local, y = O3.Mean)) +
  scale_y_continuous(name = "O3") +
  scale_x_date(date_breaks = "2 year", name = "Date", date_labels  = "%Y") + 
  ggtitle("Weekly O3 mean") + 
  theme_bw()
```


---

- If the data is grouped by week, ignoring the missing values when aggregating, it'll have 33 missing observations.
- Around 4.3% missing data.

```{r}

temp <- weekData[!is.na(weekData$O3.Mean), ]
skips <- temp$Date.Local - lag(temp$Date.Local)
skips <- skips[skips > 7][-1]/7

ggplot(data = as.data.frame(skips)) +
  geom_bar(aes(skips), fill = "blue") +
  scale_y_continuous(name = "Count") +
  scale_x_continuous(name = "Quantity of weeks", breaks = seq(0, 16, 2)) +
  ggtitle("Weeks missing in sequence") +
  theme_bw()
```

---

```{r}
imputedWeek <- cbind(weekData)
imputedWeek <- kNN(imputedWeek, variable = "O3.Mean", k = 4, 
                   numFun = weighted.mean, weightDist =  TRUE)

ggplot() +
  geom_point(data = imputedWeek[is.na(weekData$O3.Mean),], aes(x = Date.Local, y = O3.Mean), colour = 'red') +
  geom_point(data = weekData, aes(x = Date.Local, y = O3.Mean), colour = "black", alpha = 0.4) +
  scale_x_date(name ="Date", date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(name = "O3") + 
  ggtitle("Weekly Real vs Imputed data") +
  theme_bw()
```

---

- It has the same problem when the sequence of missing data is to big.
- Again, if there is more than 5 missing weeks, it will be used k = 16, if it's less, it'll be k = 4.

```{r}
imputedWeek[(imputedWeek$Date.Local > "2011-01-01") 
            & (imputedWeek$Date.Local < "2012-01-01") 
            & (is.na(weekData$O3.Mean)),"O3.Mean"] <- NA

imputedWeek <- kNN(imputedWeek, variable = "O3.Mean", k = 16, 
                   numFun = weighted.mean, weightDist =  TRUE)

ggplot() +
  geom_point(data = imputedWeek[is.na(weekData$O3.Mean), ], aes(x = Date.Local, y = O3.Mean), colour = 'red') +
  geom_point(data = weekData, aes(x = Date.Local, y = O3.Mean), alpha = 0.2, fill = "black") +
  scale_y_continuous(name = "O3") +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y", name = "Date")+
  ggtitle("Weekly Real vs Imputed data") +
  theme_bw()
```

---

```{r}
ggplot() +
  geom_point(data = imputedWeek[is.na(weekData$O3.Mean), ], aes(x = Date.Local, y = O3.Mean), colour = 'red') +
  geom_point(data = weekData, aes(x = Date.Local, y = O3.Mean), alpha = 0.2, fill = "black") +
  scale_y_continuous(name = "O3") +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y", name = "Date", limits = c(as.Date("2011-01-01"), as.Date("2012-01-01"))) +
  ggtitle("Weekly Real vs Imputed data") +
  theme_bw()
```
