---
title: "Intervention analysis"
author: "Giovani Valdrighi, Vitória Guardieiro"
date: "11/12/2020"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE, message = FALSE)
library(tseries)
library(zoo)
library(ggplot2)
library(forecast)
library(dplyr)
library(lmtest)
library(pander)
library(imager)
library(tsoutliers)
library(TSA)
```

# Total Business Sales FRED

- Data from 01/2002 to 12/2012, intervention on 07/2008. The period of 01/2013 to 12/2014 will be used as model validation.
- In 2008 we have a international banking crisis, strongly affecting sales on the US.

---

```{r}
data <- read.csv('TOTBUSSMNSA.csv')
data <- transform(data, DATE = as.Date(DATE))
title = 'Total Business Sales FRED (FED St. Louis Economics Research)'
#data before intervention
pre_intervention <- data %>% filter(DATE < as.Date("2008-09-01"))
fred_FIT <- data %>% filter(DATE < as.Date("2013-01-01"))
FRED_pre <- ts(pre_intervention$TOTBUSSMNSA,
           start = c(2002, 1),
           frequency = 12)


#fit data
FRED <- ts(fred_FIT$TOTBUSSMNSA, start = c(2002, 1), frequency = 12)

#all data
FRED_all <- ts(data$TOTBUSSMNSA, start = c(2002, 1), frequency = 12)
#check if variance is the same
autoplot(FRED, main = title) +
  scale_x_continuous(breaks = scales::extended_breaks(10))
```

---

# Variance stabilization

- With BoxCox transformation, the lambda is 0.1370143.

```{r}

#series need stabilization
lambda <- BoxCox.lambda(FRED_all)

#data after stabilization
autoplot(BoxCox(FRED, lambda), main = title) +
  scale_x_continuous(breaks = scales::extended_breaks(10))


```

---

# SARIMA pre-intervention

- Model with data until 07/2008. The plots and the models will use the transformed series.

```{r}
autoplot(BoxCox(FRED_pre, lambda), main = title) + 
  scale_x_continuous(breaks = scales::extended_breaks(10))
```

---

## Check differences

- To identify parameters $d$ and $D$ for the SARIMA model.

```{r}
#checking derivation

par(mfrow = c(2, 2))
FRED_pre %>% BoxCox(lambda = lambda) %>% plot(main = "Series")
FRED_pre %>% BoxCox(lambda = lambda) %>% diff() %>% plot(main = "diff Series")
FRED_pre %>% BoxCox(lambda = lambda) %>% diff() %>% diff() %>% plot(main = " diff diff Series")
FRED_pre %>% BoxCox(lambda = lambda) %>% diff() %>% diff(lag = 12) %>% plot(main = "diff diff(12) Series")

```

---

- Use Augmented Dickey Fuller test to verify if there is trend. 
- ADF test p-values:
  - Original series: 0.01
  - Diff() series: 0.0206
  - Diff() Diff() series: 0.01
  - Diff() Diff(12) series: 0.01
- We will be using $d= 1$ and $D = 0$ or $D= 1$.

---

## ACF and PACF

- Plot of ACF and PACF for model to identify parameters $p, q, P, Q$.

```{r}

#checking the acfs
par(mfrow = c(2, 2))
FRED_pre %>% BoxCox(lambda = lambda) %>% diff() %>% acf(main = "diff Series", lag.max = 48)
FRED_pre %>% BoxCox(lambda = lambda) %>% diff(lag = 12) %>% acf(main = "diff(12) Series", lag.max =  48)

#checking the pacfs
#par(mfrow = c(1, 2))
FRED_pre %>% BoxCox(lambda = lambda) %>% diff() %>% pacf(main = "diff Series", lag.max = 48)
FRED_pre %>% BoxCox(lambda = lambda) %>%diff(lag = 12) %>% pacf(main = "diff(12) Series", lag.max =  48)
```

---

- Both non-seasonal and seasonal ACFs decrease slowly, indicating an auto regressive model.
- For the non seasonal PACFs, we have the first two significant lags and for the seasonal we have the first three significant lags.
- Going to test the following models:
  - SARIMA(2, 1, 1)(3, 0, 1)
  - SARIMA(2, 1, 1)(2, 0, 1)
  - SARIMA(2, 1, 1)(2, 1, 1)
  

---

## SARIMA(2, 1, 1)(3,0 ,1)

```{r }
model1 <- Arima(FRED_pre, order = c(2, 1,1), seasonal = c(3, 0, 1), lambda = lambda)
myjpg <- load.image('image1.jpg')
plot(myjpg,  xaxt='n', ann=FALSE, axes = FALSE)
```

---

```{r}
checkresiduals(model1, test = FALSE)
```

---

## SARIMA(2, 1, 1)(2,0 ,1)

```{r }
model2 <- Arima(FRED_pre, order = c(2, 1,1), seasonal = c(2, 0, 1), lambda = lambda)
myjpg <- load.image('image2.jpg')
plot(myjpg,  xaxt='n', ann=FALSE, axes = FALSE)
```

---

```{r}
checkresiduals(model2, test = FALSE)
```

---

## SARIMA(2, 1, 1)(2,1 ,1)

```{r }
model3 <- Arima(FRED_pre, order = c(2, 1,1), seasonal = c(2, 1, 1), lambda = lambda)
myjpg <- load.image('image3.jpg')
plot(myjpg,  xaxt='n', ann=FALSE, axes = FALSE)
```

---

```{r}
checkresiduals(model3, test = FALSE)
```


---

- The model with lowest AIC is the SARIMA(2, 1, 1)(2, 1, 1).

```{r}
plot(FRED_pre, lwd = 1.5, main = "FRED pre intervention Real x Predicted")
lines(model3$fitted, col = 'red', lwd = 1.5)
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```

---

# Forecast of pre intervention

- We now look on how our model says that the series should be without the intervention, predicting the next six years after july 2008.

```{r}
plot(forecast(model3, h = 48, lambda =lambda, biasadj = TRUE), main = "FRED series x Forecast of pre-intervention model")
lines(FRED)
legend("topleft", legend=c("Forecast mean",  "90%", "95%", "Real values"), col=c('blue','#596DD5', "#D5DBFF" ,'black' ), lty = 1:1, cex=0.8)
```

# Intervention modeling

## Permanent constant effect

- For the first intervention model we will define $h_t = I(t > jul2008)\delta_0$, so there will be a permanent effect equal to $\delta_0$ after july of 2008.
- The estimated $\delta_0$ value is $0.0759$ with s.e. $0.1385$, so the interval includes $0$.

```{r }
#julh of 2008 is the 79 row
permanent_const_eff <-  1*(seq_along(fred_FIT$DATE) >= 79)
model4 <- Arima(FRED, order = c(2, 1, 1), seasonal = c(2, 1, 1), xreg = permanent_const_eff, lambda = lambda)
myjpg <- load.image('image4.jpg')
plot(myjpg,  xaxt='n', ann=FALSE, axes = FALSE)
```

---

```{r}
checkresiduals(model4, test = FALSE)
```

---

```{r}
plot(FRED, lwd = 1.5, main = "FRED Real x Predicted with permanent constant effect")
lines(model4$fitted, col = 'red', lwd = 1.5)
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```

---

## Temporary constant effect

- Now for the intervention model we will define $h_t = I(t = jul2008)\delta_0$, so there will be a temporary effect equal to $\delta_0$ on july of 2008.
- The estimated $\delta_0$ value is $0.1835$ with s.e. $0.1064$, so the interval doesn't include $0$.

```{r }
#julh of 2008 is the 79 row
temporary_const_eff <-  1*(seq_along(fred_FIT$DATE) == 79)
model5 <- Arima(FRED, order = c(2, 1, 1), seasonal = c(2, 1, 1), xreg = temporary_const_eff, lambda = lambda)
myjpg <- load.image('image5.jpg')
plot(myjpg,  xaxt='n', ann=FALSE, axes = FALSE)
```

---

```{r}
coeftest(model5)
```

---

```{r}
checkresiduals(model5, test = FALSE)
```

---


```{r}
plot(FRED, lwd = 1.5, main = "FRED Real x Predicted with temporary constant effect")
lines(model5$fitted, col = 'red', lwd = 1.5)
legend("topleft", legend=c("Predictions", "Real values"), col=c('red','black'), lty = 1:1, cex=0.8)
```

---

# Predictive model

```{r}
# train data
train_data <- data %>% filter(DATE < as.Date("2013-01-01"))
FRED_train <- ts(train_data$TOTBUSSMNSA,
           start = c(2002, 1),
           frequency = 12)

train_data_12 <- data %>% filter(DATE < as.Date("2014-01-01"))
FRED_train_12 <- ts(train_data_12$TOTBUSSMNSA,
           start = c(2002, 1),
           frequency = 12)
```


```{r}
interval_const_eff_train <-  1*((seq_along(train_data$DATE) >= 79 & seq_along(train_data$DATE) <= (79 + 3)))
model6_train <- Arima(FRED_train, order = c(2, 1, 1), seasonal = c(2, 1, 1), xreg = interval_const_eff_train, lambda = lambda)
summary(model6_train)
```

---

## Predict 12 months


```{r}
plot(forecast(model6_train, lambda =lambda, xreg = rep(1, each=12), biasadj = TRUE), main = "FRED series x Forecast with intervention model")
lines(FRED_train_12)
legend("topleft", legend=c("Forecast mean",  "90%", "95%", "Real values"), col=c('blue','#596DD5', "#D5DBFF" ,'black' ), lty = 1:1, cex=0.8)
```

---

## Predict 24 months

```{r}
plot(forecast(model6_train, lambda =lambda, xreg = rep(1, each=24), biasadj = TRUE), main = "FRED series x Forecast with intervention model")
lines(FRED)
legend("topleft", legend=c("Forecast mean",  "90%", "95%", "Real values"), col=c('blue','#596DD5', "#D5DBFF" ,'black' ), lty = 1:1, cex=0.8)
```

